Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2019-12-19 analysis/paper/paper.Rmd
move some code for shape to script file
all code runs

```{r tidy-data, eval=FALSE}
# Once we have the jpgs, start from here
library(tidyverse)

# get list of jpgs only
jpgs <- list.files("analysis/data/raw_data/outline_jpg", pattern = ".jpg", full.names = TRUE)

# join file names to phase
pottery_data <- readr::read_csv("analysis/data/raw_data/KWL_Ceramic_final2018_with_phases.csv")
pottery_data <- pottery_data %>% 
  mutate(filename = str_glue('fill_{filename}.svg_output.jpg'))

@@

```{r jpg-coordinates, eval = FALSE}
# start shape exploration
library(Momocs)
#  it freezes on the empty ones...
jpgs_imported <- import_jpg(jpgs, threshold = 0.5)

@@

```{r EFA-analysis}
jpgs_imported_coo <- read_rds(here::here("analysis/data/derived_data/jpgs_imported_coo.rds"))

# take a look when they stack together
stack(jpgs_imported_coo)

# fit the x and y coordinates separately
coo_oscillo(jpgs_imported_coo[1], "efourier")

# default is 13 harmanics that gather 99% of the harmonic power 
op <- efourier(jpgs_imported_coo, norm = FALSE)

@@

# mean shape, per group
pot.ms <- mshapes(op, 2)
pot.ms$shp$`Post-European` <- pot.ms$shp$`Post-European` %>% coo_plot(border="blue")
`Post-European` <- pot.ms$shp$`Post-European` %>% coo_draw(border="blue")
`Pre-European` <- pot.ms$shp$`Pre-European`  %>% coo_draw(border="red")
`Chinese` <- pot.ms$shp$`Chinese`  %>% coo_draw(border="green")
legend("topright", 
       lwd=1,
       col=c("red", "blue", "green"), 
       legend=c( "Pre-European", "Post-European", "Chinese"))

# from mshapes directly
jpgs_imported_coo %>% efourier() %>% mshapes(~phase) %>% plot_mshapes()

@@

# Isodeformation lines
tps_iso(`Pre-European`, `Post-European`,  shp.lwd = c(2, 2), shp.border = c("red", "blue"),
        amp = 0.5, iso.nb = 10000, iso.levels = 12, legend.text = c("Pre-contact", "Post-contact"),
        palette = col_summer, grid = F)
```

@@

len <- length(op.p$x[,1])
# plot PC1-3 distribution by phase
pc_plot <- data_frame(pc = c(op.p$x[,1],
                             op.p$x[,2],
                             op.p$x[,3]),
                      phase = rep(op.p$fac$phase, 3),
                        pcn = str_glue('PC{sort(rep(1:3, len))}'))

@@

ggplot(pc_plot,
       aes(phase,
           pc)) +
  geom_violin() +
  ggforce::geom_sina() +
  xlab("") +
  facet_wrap(~ pcn, ncol = 1) +
  theme_minimal()
ggsave("index/data/pottery-shape/pcs-boxplot-for-shape-variation.png",
       h = 10, w = 5)

@@

# statistical test
pc_plot_tests <- 
pc_plot %>% 
  nest_legacy(pc, phase) %>% 
  mutate(var_test = purrr::map(data, ~broom::tidy(var.test(.x$pc, .x$phase)))) %>% 
  unnest_legacy(var_test)

@@

ggsave("index/data/pottery-shape/cv-test-on-pcs.png", h = 3, w = 12)

>>>>

```{r tidy-data}
# Once we have the jpgs, start from here
library(tidyverse)
library(here)

# get list of jpgs only
jpgs <- list.files(here("analysis/data/raw_data/outline_jpg"), pattern = ".jpg", full.names = TRUE)

# join file names to phase
pottery_data <- readr::read_csv(here("analysis/data/raw_data/KWL_Ceramic_final2018_with_phases.csv"))
pottery_data <- pottery_data %>% 
  mutate(filename = str_glue('fill_{filename}.svg_output.jpg'))

@@

```{r jpg-coordinates, eval=FALSE}
# start shape exploration
library(Momocs)
#  it freezes on the empty ones...
jpgs_imported <- import_jpg(jpgs, threshold = 0.5)

@@

```{r EFA-analysis}
library(Momocs)
jpgs_imported_coo <- readr::read_rds(here::here("analysis/data/derived_data/jpgs_imported_coo.rds"))

# default is 13 harmanics that gather 99% of the harmonic power 
op <- efourier(jpgs_imported_coo, norm = FALSE, 13)

@@

# mean shape, per group
pot.ms <- MSHAPES(op, 'phase')
mshp <- pot.ms$shp
names(mshp) <- c("chi", "poe", "pre")

`Post-European` <- mshp$poe %T>%  coo_plot( border="blue")
`Pre-European` <-  mshp$pre  %T>% coo_plot(border="red", plot.new=FALSE)
`Chinese` <-       mshp$chi  %T>% coo_plot(border="green", plot.new=FALSE)
legend("topright", 
       lwd=1,
       col=c("red", "blue", "green"), 
       legend=c( "Pre-European", "Post-European", "Chinese"))

# from mshapes directly, nice but the labels spill outside of the plot area, not clear how to fix
op %>% MSHAPES(~phase) %>% plot_MSHAPES()

@@

# Isodeformation lines
png(here::here("analysis/figures/tps-iso.png"), res = 600, w = 1200, h = 1800)
tps_iso(`Pre-European`, `Post-European`,  shp.lwd = c(2, 2), shp.border = c("red", "blue"),
        amp = 0.5, iso.nb = 10000, iso.levels = 12, legend.text = c("Pre-contact", "Post-contact"),
        palette = col_summer, grid = F)
dev.off()
```

@@

len <- length(op.p$x[,1])
# plot PC1-3 distribution by phase
pc_plot <- tibble(pc = c(op.p$x[,1],
                             op.p$x[,2],
                             op.p$x[,3]),
                      phase = rep(op.p$fac$phase, 3),
                        pcn = str_glue('PC{sort(rep(1:3, len))}'))

@@

ggplot(pc_plot,
       aes(phase,
           pc)) +
  geom_violin() +
  ggforce::geom_sina() +
  xlab("") +
  facet_wrap(~ pcn, ncol = 1) +
  theme_minimal()
ggsave(here("analysis/figures/pcs-boxplot-for-shape-variation.png"),
       h = 10, w = 5)

@@

# statistical test

@@

ggsave(here("analysis/figures/cv-test-on-pcs.png"), h = 3, w = 12)
